---
title: "Quick Start"
author: "Michael Koohafkan"
date: "2019-04-15"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick Start}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This document gets you up and running with `hdfqlr`, an R interface to
[HDFql](http://www.hdfql.com/). In order to use this package, you will 
need to [download HDFql](http://www.hdfql.com/#download) for your 
system.

If you are going to be using HDFql regularly, it's a good idea to set
a default HDFql directory for use with `hdfqlr`. You can do this 
by either
 - Setting the R options `hdfqlr.dir` in your .Rprofile file.
 - Setting the system environment variable `HDFQL-DIR`.
If either of these settings exist, the HDFql library will be loaded
on package start.

```{r, echo = TRUE}
library(hdfqlr)
```

Otherwise, you can load the HDFql library after package start with
`hql_load()`:

```{r eval = FALSE}
hql_load('path/to/hdfql-x.x.x')
```

The `hdfqlr` package relies on the R wrapper provided by HDFql. 
Functions exported by the package are prefixed with `hql_` to 
make it easy to differentiate them from the functions provided 
by the wrapper, which are prefixed with `HDFQL_` (for constants)
or `hdfql_` (for functions). The contents of the HDFql wrapper
are imported into the environment `HDFql.constants`. If you need
to directly use the HDFql wrapper functions in an interactive 
session, you can access them using `with` or `attach` the 
environment to the search path.

```{r, echo = TRUE}
# using with
with(HDFql.constants,
	HDFQL_VERSION
)
# or attach the environment
attach(HDFql.constants)
HDFQL_VERSION
detach(HDFql.constants)
```

The `hdfqlr` package is designed for simple read and write use
cases, i.e. reading and writing HDF datasets and attributes. 
Creation of files, groups, and datasets is done "behind the scenes" 
in the process of writing objects to the HDF file. The following
example creates a new HDF file and writes a matrix to the file. The
dataset is then read back in and compared to the original R object.

```{r}
file = tempfile(fileext = ".h5")
values = matrix(rnorm(1000), nrow = 50)
hql_write_dataset(values, file, "dataset0")
identical(values, hql_read_dataset(file, "dataset0"))
```

Any attributes of the R object (other than `dim`) are also
written to the dataset. To write an attributes (or list of 
attributes) to the root of the HDF file or to a group, use
`hql_write_attribute` (or `hql_write_all_attributes`). 

The package does not currently support editing existing datasets
or attributes; this is instead accomplished by reading the dataset
or attribute to an R object, modifying it, and writing it back to
the HDF file using the argument `overwrite = TRUE`. Datasets, 
attributes, and groups can be removed from an HDF file by writing
`NULL` to the dataset or attribute path.

```{r}
hql_write_dataset(NULL, file, "dataset0", overwrite = TRUE)
```

